Test susieI using chr17 to 22, multiple batch

### 2020-10-14 analysis/description.Rmd
wg susieI

```{r param}
n <- 20000 # number of samples
p <- 86000 # number of SNPs
```

The genotype data we used is from UKB biobank, randomly selecting n=`r as.character(n)` samples. We use SNP genotype data from chr 17 to chr 22 combined. These genomic regions represents 12.5% of the genome. SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back. This ends up with p = `p as.charater(p)` SNPs.

@@

The one we used in this analysis is GTEx Adipose tissue v7 dataset. This dataset contains ~ 380 samples. FUSION/TWAS were used train expression model and we used their lasso results. There are around 1700 genes with expression

>>>>

```{r parameters}
n <- 40000 # number of samples
p <- 656321 # number of SNPs
J <- 8021 # number of genes
```

The genotype data we used is from UKB biobank, randomly selecting n = `r as.character(n)` samples. We use SNP genotype data from chr 1 to chr 22 combined from UKB. SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back. This ends up with p = `p as.charater(p)` SNPs.

@@

The one we used in this analysis is GTEx Adipose tissue v7 dataset. This dataset contains ~ 380 samples. [FUSION/TWAS](http://gusevlab.org/projects/fusion/) were used to train expression model and we used their lasso results. SNPs included in eQTL anlaysis are restricted to cis-locus 500kb on either side of the gene boundary. eQTLs are defined as SNPs with abs(effectize) > 1e-8 in lasso results.

2. Impute expression.

We impute gene expression for our genotype data using expression models obtained from step 1. There are `r as.character(J)` genes with expression model from chr17 to chr22. We imputed expression from genotypes using the expression predictors.

3. Define and select regions

Next, the analysis is done at the “region” level, which is 500kb bins along the genome. Each bin would contain all the SNPs, as well as all the genes in that bin. We are exploring several ways to select regions that contain true signals, e.g. based on regional sum of mr.ash PIP for genes/SNPs, region smallest TWAS p value for gene/SNPs, or regional bayes factors, etc.

4. Run susie iteratively
We then run susie for each of these regions. So the features of SuSiE are: SNPs and "genes" (not cis-eQTLs of that gene). We use the same prior for all SNPs and another prior for all "genes" when running SUSIE. In some settings, we also run SUSIE with null weight, which is calculated as `1- prior.SNP * n.SNP - prior.gene * n.gene`. We obtain the PIP for SNPs and gene in the region. After we run susie for all regions (one iteration), we take the average of all SNP PIPs as the prior of SNPs for the next iteration and similarly for the prior for genes.

5. We obtain PIP for genes from the last iteration as results.

### 2020-10-14 analysis/simulation-susieI-ukbchr17to22B-gtex.adipose.Rmd
wg susieI

---
title: "Test susieI using chr17 to 22"
output: html_document
---

@@

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200915/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20200915/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20200915/"
tag <- '20200915-1-1'
Niter <- 30
```

# Regional minimum p value distribution

* minimum TWAS p value of genes for each region:

```{r rpip twas, fig.height=6, fig.width=6}
simtag <- "20200915-1-1"
rpipf <- paste0(outputdir, simtag, ".pgenerprank.txt")

a <- read.table(rpipf, header = T)
a.c <- a[a$nCausal > 0, ]
ax <- pretty(0:max(-log10(a$rpmin)), n = 30)

par(mfrow=c(2,1))
h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
```

* minimum TWAS p value of genes and SNPs for each region:

```{r rpip twas2, fig.height=6, fig.width=6}
rpipf <- paste0(outputdir, simtag, ".p2rprank.txt")

a <- read.table(rpipf, header = T)
a.c <- a[a$nCausal > 0, ]
ax <- pretty(0:max(-log10(a$rpmin)), n = 30)

par(mfrow=c(2,1))
h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
```

>>>>

---
title: "Test susieI using chr17 to 22, multiple batch"
output: html_document
---

@@

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200915/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20200915/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20200915/"
tag <- '20200915-1-1'
Niter <- 30
```
# Summary

This result is problematic. The artificial perfectly correlated LD leads to overestimated pi1.
